---
title: "Assignment 4 Analysis"
author: "Josh Burroughs"
date: "February 28, 2016"
output: html_document
---

```{r}
library(data.table)
library(bit64)
```

# San Francisco
Downloaded from https://data.sfgov.org/view/vsk2-um2x 
```{r}
sf <- fread("data/sf/SFPD_Incidents_-_from_1_January_2003.csv")
setnames(sf,make.names(names(sf)))
sf.factor.cols <- c('Category','Descript','DayOfWeek','Resolution')
sf[,sf.factor.cols := lapply(.SD,factor),.SDcols = sf.factor.cols, with=F]
sf[,Date := as.POSIXct(Date,format='%m/%d/%Y',tz='PST')]
t_to_secs <- function(time) {
  lt <- as.POSIXlt(time,format='%M:%S')
  lt$min * 60 + lt$sec
}
sf[,DateTime := Date + t_to_secs(Time)]
```

```{r}
library(ggplot2)
with(sf,smoothScatter(x=X,y=Y))
```
Data outside the main geographical area:
```{r}
boxplot(sf$X)
title("X")
boxplot(sf$Y)
title("Y")
max.loc = sf[,sapply(.SD,max),.SDcols=c('X','Y')]
```
Looks like points at `r max.loc` are bogus
```{r}
sf[X == max.loc[['X']] & Y == max.loc[['Y']],`:=`(X=NA,Y=NA)]
with(sf,smoothScatter(x=X,y=Y))
```

Looking for murders:
```{R}
sf[grep('HOMICIDE|MURDER|SLAUGHTER',Descript),unique(Descript)]
sf[grep('ASSAULT',Descript),unique(Descript)]
```
Only attempted homicide...
```{r}
sf[grep('GRAND THEFT',Descript),unique(Descript)]
```

# Raleigh

Downloaded from https://data.raleighnc.gov/Police/Police-Incident-Data-from-Jan-1-2005-Master-File/csw9-dd5k
```{r}
ral <- fread("data/raleigh/Police_Incident_Data_from_Jan_1_2005_-_Master_File.csv")
setnames(ral,make.names(names(ral)))
ral.factor.cols <- c('LCR','DISTRICT')
ral[,ral.factor.cols := lapply(.SD,factor),.SDcols = ral.factor.cols, with=F]
#split the Location column apart:
ral[,c('Y','X') := lapply(tstrsplit(gsub('[()]','',LOCATION),','),as.numeric)]
ral[,INC.DATETIME := as.POSIXct(INC.DATETIME,format="%m/%d/%Y %I:%M:%S %p",tz='EST')]
```

```{r}
with(ral,smoothScatter(X,Y))
```

Incident Code:
```{r}
ralgrep <- function(term) unique(ral[grep(term,LCR.DESC,ignore.case = T),.(LCR,LCR.DESC)])
ralgrep('murder')
ralgrep('attempt')
ralgrep('theft')
ralgrep('larc')
ralgrep('assault')
```

From here on, we'll analyze GTA and combined assault incidents.  Assign category names to the crimes we're interested in:
```{r} 
ral[grep('assault',LCR.DESC,ignore.case = T), Cat := 'ASSAULT']
ral[LCR %in% c('40A','71','72','73'), Cat := 'GTA']
sf[grep('AUTO, GRAND THEFT|GRAND THEFT AUTO',Descript), Cat := 'GTA']
sf[grep('ASSAULT',Descript), Cat := 'ASSAULT']
```
Join the two data sets:
```{r}
crimes <- rbind(
  ral[,.(City='Raleigh',DateTime = INC.DATETIME,X,Y,Category=Cat)],
  sf[,.(City='San Francisco',DateTime,X,Y,Category=Cat)])[!is.na(Category),]
factor.cols <- c('City','Category')
crimes[,factor.cols := lapply(.SD,as.factor),.SDcols = factor.cols,with=F]
```
Only use dates where both cities have data:
```{r}
max.t <- min(crimes[,.(d = max(DateTime)),by=City][,d])
min.t <- max(crimes[,.(d = min(DateTime)),by=City][,d])
crimes <- crimes[DateTime > min.t & DateTime < max.t]
```

We need to normalize by population to compare the 2 cities.
Census population estimate for 2000-2009, downloaded from 
https://www.census.gov/popest/data/cities/totals/2009/ on 2/28/2016:
```{r}
pop.2009 <- fread("grep -E '^[0-9]' data/SUB-EST2009-01.csv",skip=3)
setnames(pop.2009,c('Rank','City','State',2009:2000,'Base','Census'))
pop.2014 <- fread("data/PEP_2014_PEPANNRSIP.US12A_with_ann.csv")
setnames(pop.2014,c(make.names(names(pop.2014)[1:8]),'Census','Base',2010:2014))
p.2009 <- melt(pop.2009,1:3,4:13,variable.name='Year',value.name="Population")[grep('San Francisco|Raleigh',City),.(City = gsub(' city','',City),Year,Population)]

p.2014 <- melt(pop.2014[grep('San Francisco|Raleigh',GC_RANK.display.label)],c('GC_RANK.display.label'),11:15,variable.name='Year',value.name='Population')[,.(City = gsub('United States - | city,.*','',GC_RANK.display.label),Year,Population)][grep('South',City,invert = T)]

predpop <- function(p,year) {
  m <- lm(Population ~ Year,p)
  predict(m,2015)
}
population <- rbind(p.2009,p.2014)
population[,Year := as.numeric(as.character(Year))]
population[,Population := as.numeric(gsub(',','',Population))]

#use lm to project 2015 populations
p.2015 <- population[,.(Year=2015,Population=predict(lm(Population~Year),newdata=data.frame(Year=2015))),by=City]
population <- rbind(population,p.2015)
#population <- population[order(City,Year)]
```

Now to join the crime data with population data
```{r}
crimes[,Year := as.numeric(format(DateTime,'%Y'))]
crimes <- crimes[Year < 2016]
crimes <- merge(crimes,population,by=c('City','Year'))
```

```{r}
annual.crime <- crimes[,.(Per.100K = .N/(Population/100000)),by=.(City,Category,Year)]

ggplot(annual.crime,aes(x=Year,y=Per.100K,fill=City))+geom_bar(stat='identity',position='dodge') + labs(y='Incidents Per 100K Pop') + facet_wrap(~ Category,scales='free_y',ncol=1)

ggplot(annual.crime,aes(x=Year,y=Per.100K,fill=Category))+geom_bar(stat='identity',position='dodge') + labs(y='Incidents Per 100K Pop') + facet_wrap(~ City,ncol=1)
```

```{r}
library(ggmap)
city.centers <- crimes[!is.na(X) & !is.na(Y),lapply(.SD,function(V) (max(V) - min(V))/2 + min(V)),by=City,.SDcols=c('Y','X')][,.(City,lon=X,lat=Y)]
center.sf <- city.centers[City=='San Francisco']
center.ral <- city.centers[City=='Raleigh']
map.sf <- get_map(location = c(center.sf$lon,center.sf$lat),zoom=12)
map.ral <- get_map(location = c(center.ral$lon,center.ral$lat),zoom=12)
blankout <- theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      legend.position="none",
      panel.background=element_blank(),
      panel.border=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      plot.background=element_blank())
map_progression <- function(m,city,cat) {
  print(ggmap(m) + geom_density_2d(aes(x = X, y = Y,color=Category),data=crimes[City==city & Category == cat]) + facet_wrap(~ Year,nrow=2) + blankout + labs(title=paste(cat,"in",city)))
}
map_progression(map.sf,'San Francisco','GTA')
map_progression(map.ral,'Raleigh','GTA')
map_progression(map.sf,'San Francisco','ASSAULT')
map_progression(map.ral,'Raleigh','ASSAULT')
```



